---
title: "Socioeconomic Drivers of Health"
subtitle: "Income, Anxiety, and Substance Use"
author: "STA220 Project Team"
format: 
  revealjs:
    theme: night
    transition: slide
    slide-number: true
    scrollable: true
---

## Research Questions

1.  Is **income** associated with the prevalence of **severe anxiety**?
2.  Is **income** associated with **substance use disorders**?
3.  Do these associations differ by **age** or **gender**?

## Data Overview

We linked two primary datasets for this analysis:
- **Patients Dataset**: Demographic info (Income, Age, Gender).
- **Conditions Dataset**: Clinical records (SNOMED codes for Anxiety and Substance Use).

```{r setup, include=FALSE}
library(targets)
library(ggplot2)
library(gt)
```

## Population Demographics

*A look at our cohort distribution across income quartiles.*

```{r}
#| echo: false
tar_read(tab_demo)
```

## Income vs. Condition Prevalence

*Does earning more correlate with lower health risks?*

```{r}
#| echo: false
#| fig-align: center
tar_read(fig_prev)
```

## The Role of Gender & Age

*Exploring if the income-health relationship is universal.*

```{r}
#| echo: false
#| fig-align: center
tar_read(fig_interact)
```
