---
title: "Socioeconomic Drivers of Health"
subtitle: "Income, Anxiety, and Substance Use"
author: "Amal Ben Zalouh & Roshinie C. De Silva"
format: 
  revealjs:
    theme: night
    transition: slide
    slide-number: true
    scrollable: true
---

## Research Questions

1. Is **income** associated with the prevalence of **severe anxiety**?
2. Is **income** associated with **substance use disorders**?
3. Do these associations differ by **age** or **gender**?

## Data Overview

We linked two primary datasets for this analysis:

- **Patients Dataset**: Demographic info (Income, Age, Gender)
- **Conditions Dataset**: Clinical records (SNOMED codes for Anxiety and Substance Use)

**Data cleaning:**
- Removed deceased patients
- Removed impossible ages
- Removed extreme income outlier ($20)
- Kept living adults aged 18+
- Log-transformed income to address right skew
```{r setup, include=FALSE}
library(targets)
library(ggplot2)
library(gt)
library(knitr)
```

## Population Demographics

*A look at our cohort distribution across age groups and gender.*
```{r}
#| echo: false
tar_read(tab_demo)
```

## Income vs. Condition Prevalence

*Does earning more correlate with lower health risks?*
```{r}
#| echo: false
#| fig-align: center
tar_read(fig_prev)
```

## Statistical Results: Addiction (35 and over)

*Logistic regression — Addiction ~ Income + Gender*
```{r}
#| echo: false
tar_read(models)$addiction_35plus |>
  kable(digits = 3, col.names = c("Term", "OR", "Std Error", 
                                   "Statistic", "P-value", 
                                   "CI Low", "CI High"))
```

> Higher income is associated with **lower odds of addiction** in adults 35 and over (OR = 0.85, CI [0.80-0.91])

## Statistical Results: Anxiety (All Adults)

*Logistic regression — Anxiety ~ Gender + Age Group*
```{r}
#| echo: false
#| fig-align: center
tar_read(fig_interact)
```
